Smart Spreading Factor Assignment for LoRaWANs

### 2019-02-26 paper.tex
revert omitted lora details

Where, $R_{b}$ is data rate in bit/sec, SF is spreading factor $SF \in \{7,..,12\}$, CR is error correction code rate $CR \in \{1,..,4\}$ and $BW$ is bandwidth in Hertz \cite{AN1200.22}.

@@

\section{Simulation Environment} \label{Simulation Environment}
\par A discrete event simulator is implemented in Python to study the effects of different SF strategies in LoRaWANs. LoRaWAN SF simulation tool source code is available at \cite{simlorasf}. Simulation tool supports custom LoRaWAN topologies as well as randomly generated LoRaWAN topologies. Simulator can generate uniformly distributed circular shape network topology with following input parameters: radius in meter (m), number of nodes and number of gateways. Global simulation input parameters are simulation duration in second (s), packet size in Byte (B), packet generation rate in packets per second (pac/sec) and SF assignment method. With these inputs, the simulator produces total number of generated packets, number of successfully received packets, number of interfered packets, number of under sensitivity packets, network packet delivery ratio percentage (PDR), network throughput in bits per second (bit/sec) and total transmit energy consumption in Joule (J). Simulator also produces prediction accuracy and confusion matrix for machine learning schemes.

@@

\par Where, $P^{dBm}_{RX}$ is the expected receive power at the receiver. $P^{dBm}_{TX}$ is the transmit power of the transmitter. $G^{dB}_{SYS}$ is the system gains such as transmitter and receiver antenna gains. $L^{dB}_{SYS}$ is the system losses such as transmitter and receiver line, circuit, antenna losses. $L^{dB}_{PATH}$ is the propagation path loss between transmitter and receiver antennas in open space. In the simulator, it is assumed that total of system gains $G^{dB}_{SYS}$ and system losses $L^{dB}_{SYS}$ is +7 dB.

\par In the simulator, it is assumed that nodes always select maximum allowed transmit power, which is 14 dBm for European ISM band regulation \cite{lorawan.regional.parameters}. Different channel transmissions are independent from each other. But, in this work we focus on SF orthogonality. Thus, only single channel transmissions are utilized in the simulator.

\par Receive sensitivity of a LoRa gateway for different SFs and BWs in dBm unit can be found in \cite{SX1276}. In the simulator, 125 kHz bandwidth receive sensitivities are used.

\par Free space propagation loss is calculated as in \cite{TR136.942}. In this work, it is assumed that $h$ = 15 m and $f$ = 868 MHz.

@@

\par It is assumed that there is no other technology interference in the network except LoRa interference. To exploit imperfect orthogonality of different SF transmission, simulator should calculate the effect of different SF transmissions to each other. In simulator, signal to interference plus noise ratio (SINR) threshold matrix from \cite{goursaud:hal-01231221} is used.

\par To decide if a referenced signal is interfered at receiver by an interfering signal, SINR threshold matrix is used. $T_{i,j}$ is SINR margin in dB unit between referenced signal with SF = i and interfering signal with SF = j to correctly decode the referenced signal. If there are more than one interfering signal, referenced signal must satisfy the margin for cumulative sum of all interfering signal received power for each SF \cite{7996384}. Transmission duration of a packet can be calculated by data rate $R_{b}$ and packet size $PS$. Data rate of a LoRa transmission is already expressed in Equation \ref{eq:bit_rate_sf}.

\section{Simulation Results} \label{Simulation Results}
\par For simulation results in this paper, global simulation parameters are set as follows: packet size = 60 Bytes, simulation duration = 3600 seconds, packet generation rate = 0.01 pac/sec.

>>>>

Where, $R_{b}$ is data rate in bps, SF is spreading factor $SF \in \{7,..,12\}$, CR is error correction code rate $CR \in \{1,..,4\}$ and $BW$ is bandwidth in Hertz \cite{AN1200.22}.

@@

\section{Simulation Environment} \label{Simulation Environment}
\par A discrete event simulator is implemented in Python to study the effects of different SF strategies in LoRaWANs. LoRaWAN SF simulation tool source code is available at \cite{simlorasf}. Simulation tool supports custom LoRaWAN topologies as well as randomly generated LoRaWAN topologies. Simulator can generate uniformly distributed circular shape network topology with following input parameters: radius (r) in meter (m), number of nodes and number of gateways. Global simulation input parameters are simulation duration in second (s), packet size in Byte (B), packet generation rate in packets per second (pps) and SF assignment method. With these inputs, the simulator produces total number of generated packets, number of successfully received packets, number of interfered packets, number of under sensitivity packets, network packet delivery ratio percentage (PDR), network throughput in bits per second (bps) and total transmit energy consumption in Joule (J). Simulator also produces prediction accuracy and confusion matrix for machine learning schemes.

@@

\par Where, $P^{dBm}_{RX}$ is the expected receive power at the receiver. $P^{dBm}_{TX}$ is the transmit power of the transmitter. $G^{dB}_{SYS}$ is the system gains such as transmitter and receiver antenna gains. $L^{dB}_{SYS}$ is the system losses such as transmitter and receiver line, circuit, antenna losses. $L^{dB}_{PATH}$ is the propagation path loss between transmitter and receiver antennas in open space. In the simulator, it is assumed that sum of system gains $G^{dB}_{SYS}$ and system losses $L^{dB}_{SYS}$ is +7 dB.

\begin{table}
\centering
\caption{EU863-870 ISM band LoRa default channels \cite{lorawan.regional.parameters}.}
\label{table:max_tx_power}
\begin{tabular}{|c|c|c|}
\hline
\textbf{Frequency (MHz)} & \textbf{Bandwidth (kHz)} & \textbf{Max EIRP (dBm)} \\ \hline
      868.1 &   125 &   14 \\ \hline
      868.3 &   125 &   14 \\ \hline
      868.5 &   125 &   14 \\ \hline
\end{tabular}
\end{table}

\par Maximum transmit power for European ISM band LoRa default channels can be found in Table \ref{table:max_tx_power}. In the simulator, it is assumed that nodes always select maximum allowed transmit power, which is 14 dBm. Different channel transmissions are independent from each other. But, in this work we focus on SF orthogonality. Thus, only single channel transmissions are utilized in the simulator.

\begin{table}
\centering
\caption{Gateway sensitivity for different SFs \cite{SX1276}.}
\label{table:gw_sf_sensitivity}
\begin{tabular}{|c|c|c|c|c|c|c|c|}
\hline
\multicolumn{2}{|c|}{\multirow{2}{*}{}} & \multicolumn{6}{c|}{\textbf{SF}} \\ \cline{3-8}
\multicolumn{2}{|c|}{}                  &    7 &    8 &    9 &   10 &   11 &   12 \\ \hline
\multirow{3}{*}{\textbf{BW (kHz)}}  & 125 & -123 & -126 & -129 & -132 & -133 & -136 \\ \cline{2-8}
                                    & 250 & -120 & -123 & -125 & -128 & -130 & -133 \\ \cline{2-8}
                                    & 500 & -116 & -119 & -122 & -125 & -128 & -130 \\ \hline
\end{tabular}
\end{table}

\par Receive sensitivity of a LoRa gateway for different SFs and BWs in dBm unit can be found in Table \ref{table:gw_sf_sensitivity}. In the simulator, 125 kHz bandwidth receive sensitivities are used.

\par Free space propagation loss is calculated as \cite{TR136.942}:

\begin{equation} \label{eq:propagation_loss}
\begin{split}
P^{dB}_{PATH} = 40(1 - 4 \times 10^{-3} \times h){\log_{10} R|_{km}} \\
- 18 {\log_{10} h|_{m}} + 21 {\log_{10} f|_{MHz}} + 80
\end{split}
\end{equation}

\par Where, $h$ is the gateway altitude and $f$ is the frequency of the signal. In this work, it is assumed that $h$ = 15 m and $f$ = 868 MHz. With these assumptions, propagation loss calculation become \cite{7996384}:

\begin{equation} \label{eq:propagation_loss_simplified}
P^{dB}_{PATH} = 120.5 + 37.6 {\log_{10} R|_{km}}
\end{equation}

@@

\par In this work, it is assumed that there is no other technology interference in the network except LoRa interference. To exploit imperfect orthogonality of different SF transmission, simulator should calculate the effect of different SF transmissions to each other. In simulator, signal to interference plus noise ratio (SINR) threshold matrix $T_{i,j}$ from \cite{goursaud:hal-01231221} is used.

\par To decide if a referenced signal is interfered at receiver by an interfering signal, SINR threshold matrix is used. $T_{i,j}$ is SINR margin in dB unit between referenced signal with SF = i and interfering signal with SF = j to correctly decode the referenced signal. If there are more than one interfering signal, referenced signal must satisfy the margin for cumulative sum of all interfering signal received power for each SF \cite{7996384}. SINR threshold can be calculated as:

\begin{equation} \label{eq:sinr_db}
SINR_{i,j} = \dfrac{P_{rc,0}}{\sum_{l \in I_j} P_{rc,l}}
\end{equation}

\par Where $P_{rc,0}$ is received signal power of referenced signal and $P_{rc,l}$ is received signal power of interfering signal for SF = j. If a packet with SF = i satisfies the following condition for every SF = j, then packet is survived from all interferences.

\begin{equation} \label{eq:sinr_t}
SINR_{i,j}^{dB} > T_{i,j}
\end{equation}

\par To calculate interfering power at receiver, we should consider the case where two transmissions are not perfectly overlapping. To equalize the interfering power at receiver \cite{7996384}:

\begin{equation} \label{eq:p_interference}
P_{rc,y}^{interf} = \dfrac{P_{rc,y}(t_{interf})}{t_{x}}
\end{equation}

\par Where $t_{x}$ is transmission duration of referenced signal. $t_{interf}$ is overlapping duration between referenced signal and interfering signal. Transmission duration of a packet can be calculated by data rate $R_{b}$ and packet size $PS$. Data rate of a LoRa transmission is already expressed in Equation \ref{eq:bit_rate_sf}. Duration of a LoRa transmission can be calculated as:

\begin{equation} \label{eq:transmission_duration}
t_{x} = \dfrac{8 \times PS|_{B}}{R_{b}|_{bps}}  \ s
\end{equation}

\section{Simulation Results} \label{Simulation Results}
\par For simulation results in this paper, global simulation parameters are set as follows: packet size = 60 Bytes, simulation duration = 3600 seconds, packet generation rate = 0.01 pps.
